|

### 2021-12-07 01-Chp-Background.Rmd
Fix broken cross-links

Statins are a commonly prescribed method of lipid regulation.[@collins2016] Statins inhibit the conversion of 3-hydroxy-3-methylglutaryl-coenzyme-A (HMG-CoA) into mevalonate, by competitively binding with HMG-CoA reductase (HMGCR). This conversion is a key rate-limiting step in the cholesterol biosynthesis pathway (see Figure \@ref(statin-mechanisam)), enabling statins to reduce effectively the production of LDL cholesterol.

@@

There are several other interventions that can be used to modify a persons lipid profile, each of which act via different mechanisms (Table \@ref(tab)). However, in general, these treatments are either used as adjunct (additional) treatments with statins therapy or are used in situations where statins are contra-indicated or not tolerated.

<!-- see https://www.heart.org/en/health-topics/cholesterol/prevention-and-treatment-of-high-cholesterol-hyperlipidemia/cholesterol-medications for information on the indications for treatment - could present as a table -->

@@

In this thesis, preprints are considered an important source of evidence, in contrast to previous reviews on this topic. However, as with grey literature,[@mahood2014] there are several logistical problems with carrying out systematic searches in preprint repositories. As such, to enable the inclusion of preprints in the systematic review described in Chapters \@ref(sys-rev-methods-heading), a new tool addressing these issues is presented in Chapter \@ref(sys-rev-tools).

@@

I will attempt to obtain the raw data from relevant primary studies identified by the systematic review in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading). Any data obtained will be combined with that available from the DPUK portal as part of an IPD meta-analysis in Chapter \@ref(ipd-analysis-heading). This will enable the assessment of the effect of lipids on dementia stratified by key variables such as sex.

@@

To address this limitation, this thesis seeks to develop and apply a novel quantitative triangulation approach to answer causal questions on the effect of blood lipids on dementia outcomes. All existing evidence, regardless of study design, is first identified and assessed using domain-based tools by the systematic review presented in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading). The existing evidence base is then supplemented by two primary analyses, presented in Chapters \@ref(cprd-analysis-heading) and \@ref(ipd-analysis-heading). Finally all existing and new evidence is incorporated into the quantiatative triangulation framework in Chapter \@ref(tri-heading).

@@

Using the tool described in Chapter \@ref(sys-rev-tools), I lead a "research-on-research" study to assess the concordance between the openness of data availability statements accompanying a sample of medRxiv preprints and their published counterparts.

>>>>

Statins are a commonly prescribed method of lipid regulation.[@collins2016] Statins inhibit the conversion of 3-hydroxy-3-methylglutaryl-coenzyme-A (HMG-CoA) into mevalonate, by competitively binding with HMG-CoA reductase (HMGCR). This conversion is a key rate-limiting step in the cholesterol biosynthesis pathway (see Figure \@ref(fig:statin-mechanisam)), enabling statins to reduce effectively the production of LDL cholesterol.

@@

There are several other interventions that can be used to modify a persons lipid profile, each of which act via different mechanisms. However, in general, these treatments are either used as adjunct (additional) treatments to statin therapy or are used in situations where statins are contra-indicated or not tolerated.

<!-- TODO see https://www.heart.org/en/health-topics/cholesterol/prevention-and-treatment-of-high-cholesterol-hyperlipidemia/cholesterol-medications for information on the indications for treatment - could present as a table . If using table, put reference to it in the above par -->

@@

In this thesis, preprints are considered an important source of evidence, in contrast to previous reviews on this topic. However, as with grey literature,[@mahood2014] there are several logistical problems with carrying out systematic searches in preprint repositories. As such, to enable the inclusion of preprints in the systematic review described in Chapters \@ref(sys-rev-methods-heading), a new tool addressing these issues is presented in Chapter \@ref(sys-rev-tools-heading).

@@

I will attempt to obtain the raw data from relevant primary studies identified by the systematic review in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading). Any data obtained will be combined with that available from the DPUK portal as part of an IPD meta-analysis in Chapter \@ref(ipd-heading). This will enable the assessment of the effect of lipids on dementia stratified by key variables such as sex.

@@

To address this limitation, this thesis seeks to develop and apply a novel quantitative triangulation approach to answer causal questions on the effect of blood lipids on dementia outcomes. All existing evidence, regardless of study design, is first identified and assessed using domain-based tools by the systematic review presented in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading). The existing evidence base is then supplemented by two primary analyses, presented in Chapters \@ref(cprd-analysis-heading) and \@ref(ipd-heading). Finally all existing and new evidence is incorporated into the quantiatative triangulation framework in Chapter \@ref(tri-heading).

@@

Using the tool described in Chapter \@ref(sys-rev-tools-heading), I led a "research-on-research" study to assess the concordance between the openness of data availability statements accompanying a sample of medRxiv preprints and their published counterparts.

### 2021-12-07 02-Chp-Systematic-Review-Tools.Rmd
Working version
Fix broken cross-links

Searching these preprints as part of the systematic review described in Chapter \@ref(sys-rev-heading) was a necessity, as many of the existing reviews on the topic of lipids and dementia have not considered this important source of evidence. At the time of writing, however, the bioRxiv and medRxiv websites allow only simple search queries as opposed to the complex searches containing Boolean logic operators (AND/OR/NOT) that information specialists use to query other major databases.[@bramer2018;@gusenbauer2020] Additionally, the best available extraction mechanism for obtaining references for all records returned by a search involved going through each record and downloading individual citations one-by-one. As the scale of these preprint databases increase, particularly in light of the massive expansion of the medRxiv repository as a result of COVID, this already time-consuming and error-prone method is no longer feasible.

@@

<!-- IDEA Add timeline chart as seen here: https://bookdown.org/alecri/thesis/4-methods.html#sec:dosresmeta -->

@@

I developed the tool to meet three success criteria,[@wateridge1995] influenced both by the functionality required to perform systematic searches as part of the review in Chapter \@ref(sys-rev-heading), discussion with information specialist colleagues, and an informal survey of the evidence synthesis and health librarian communities on Twitter. The criteria were as follows:

@@

Work on the `medrxivr` tool began in Summer 2019, and initially consisted of a development of set of R scripts to allow for searching medRxiv and bioRxiv as part of the systematic search outlined in Chapter \@ref(sys-rev-heading). Following interest from other researchers in using the *ad-hoc* web-scraping scripts, additional development work took place in 2019/2020, allowing for improved searching and exporting functionality and I released the initial version of the `medrxivr` R package in February 2020.

@@

In addition to being used to search systematically health-related preprint servers, as illustrated in the systematic review presented in Chapter \@ref(sys-rev-heading), `medrxivr` has other uses.

@@

While searching of the medRxiv and bioRxiv databases was crucial for the systematic review element of my thesis presented in Chapter \@ref(sys-rev-heading), there are some important limitations to note here. A key example is that the tool only searches the available metadata of preprint records (the title, abstract and keywords), rather than the full text of preprints, and so some relevant records might be missed. However, this approach echoes that used by other search platforms such as OvidSP, and while some relevant records may be missed (reduced sensitivity), limiting the search to the metadata fields prevents non-relevant records from being returned (high specificity). A key example of the reduced specificity when searching the full text, identified during development of `medrxivr`, is that a search for "dementia" would return a record where the only occurrence of this term is in the title of one of the references.[@bong2019]

@@

-   I have outlined the motivation for developing this tool in relation to this thesis - more specifically, that it was used to perform systematic and reproducible searches of key literature sources used in the comprehensive systematic review described in Chapter \@ref(sys-rev-heading).

>>>>

Searching these preprints as part of the systematic review described in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading) was a necessity, as many of the existing reviews on the topic of lipids and dementia have not considered this important source of evidence. At the time of writing, however, the bioRxiv and medRxiv websites allow only simple search queries as opposed to the complex searches containing Boolean logic operators (AND/OR/NOT) that information specialists use to query other major databases.[@bramer2018;@gusenbauer2020] Additionally, the best available extraction mechanism for obtaining references for all records returned by a search involved going through each record and downloading individual citations one-by-one. As the scale of these preprint databases increase, particularly in light of the massive expansion of the medRxiv repository as a result of COVID, this already time-consuming and error-prone method is no longer feasible.

@@

@@

I developed the tool to meet three success criteria,[@wateridge1995] influenced both by the functionality required to perform systematic searches as part of the review presented in the subsequent chapters, discussion with information specialist colleagues, and an informal survey of the evidence synthesis and health librarian communities on Twitter. The criteria were as follows:

@@

Work on the `medrxivr` tool began in Summer 2019, and initially consisted of a development of set of R scripts to allow for searching medRxiv and bioRxiv as part of the systematic search outlined in Chapter \@ref(sys-rev-methods-heading). Following interest from other researchers in using the *ad-hoc* web-scraping scripts, additional development work took place in 2019/2020, allowing for improved searching and exporting functionality and I released the initial version of the `medrxivr` R package in February 2020.

@@

In addition to being used to search systematically health-related preprint servers, as illustrated in the systematic review presented in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading), `medrxivr` has other uses.

@@

While searching of the medRxiv and bioRxiv databases was crucial for the systematic review element of my thesis, there are some important limitations to note here. A key example is that the tool only searches the available metadata of preprint records (the title, abstract and keywords), rather than the full text of preprints, and so some relevant records might be missed. However, this approach echoes that used by other search platforms such as OvidSP, and while some relevant records may be missed (reduced sensitivity), limiting the search to the metadata fields prevents non-relevant records from being returned (high specificity). A key example of the reduced specificity when searching the full text, identified during development of `medrxivr`, is that a search for "dementia" would return a record where the only occurrence of this term is in the title of one of the references.[@bong2019]

@@

-   I have outlined the motivation for developing this tool in relation to this thesis - more specifically, that it was used to perform systematic and reproducible searches of key literature sources used in the comprehensive systematic review described in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading).

### 2021-12-07 04-Chp-Systematic-Review-Results.Rmd
Fix broken links to appendices
Fix broken cross-links

Across the remaining `r num_to_text(n_dr_final)` studies, a sufficient number of results (n $\geqslant$ 3) were identified only for the total cholesterol-Alzheimer's, LDL-Alzheimer's, and total cholesterol-dementia strata. This analysis provided weak evidence for a non-linear effect of lipid levels on dementia outcomes, and Figure \@ref(fig:lipidsDoseResponse) illustrates this for the total cholesterol-Alzheimer's strata. Similar figures for the other analysed lipid-outcome strata are presented in Appendix \@ref(hold).

@@

Secondly, this review employed a structured approach to risk-of-bias assessment using a domain-based tool.  This represents an important strength of this review, as the detailed risk of bias assessments are used in inform the quantitative triangulation analysis presented in Chapter \@ref(tri-analysis-heading).

>>>>

Across the remaining `r num_to_text(n_dr_final)` studies, a sufficient number of results (n $\geqslant$ 3) were identified only for the total cholesterol-Alzheimer's, LDL-Alzheimer's, and total cholesterol-dementia strata. This analysis provided weak evidence for a non-linear effect of lipid levels on dementia outcomes, and Figure \@ref(fig:lipidsDoseResponse) illustrates this for the total cholesterol-Alzheimer's strata. Similar figures for the other analysed lipid-outcome strata are presented in Appendix \@ref(appendix-dr-fraction).

@@

Secondly, this review employed a structured approach to risk-of-bias assessment using a domain-based tool.  This represents an important strength of this review, as the detailed risk of bias assessments are used in inform the quantitative triangulation analysis presented in Chapter \@ref(tri-heading).

### 2021-12-07 05-Chp-CPRD-Analysis.Rmd
Fix broken cross-links

In this chapter, I present an analysis of a large population-based electronic health record dataset to investigate the relationship between lipid-regulating agent (LRA) use and dementia outcomes. The analysis aims to address two important limitations of the current evidence base as identified by the systematic review presented in Chapter \@ref(sis-rev-heading).

@@

This second presentation appears to be common in the existing literature on the relationship of statins and dementia, as several of the studies included in the systematic review presented in Chapter \@ref(sys-rev-heading) were identified as being at risk of immortal time bias following formal risk of bias assessment using the ROBINS-I tool (see Section \@ref(risk-of-bias-subheading)).

@@

#### Impact of dementia code lists {#comparing-code lists}

@@

This section will expand on a potential explanation for the observed results detailed above. However, as the comparison of evidence across different sources is the aim of the triangulation exercise presented in later chapters, a detailed comparison with other published literature will not be provided here, except where needed to illustrate a methodological point. For a detailed comparison of the result presented above with the existing evidence base identified by the systematic review (Chapter \@ref(sys-rev-heading)), see Chapter \@ref(tri-heading).

@@

Given the absence of vascular dementia in the published literature, as highlighted in the previous chapter (see Section \@ref(sys-rev-pub-bias)), the unexpected increase in vascular dementia risk with statin use is particularly interesting. It is possible that this finding reflects publication bias if previous research encountered similar methodological issues to this analysis, and these results did not make it into the evidence base as unexpected or assumedly incorrect results are less likely to be submitted or accepted for publication.

@@

The primary strength of this analysis is the relative size of the CPRD and length of follow-up. Having reviewed the existing literature, as identified by the systematic review in Chapter \@ref(sys-rev-heading), this analysis of `r comma(attrition[14,1])` participants is one of the largest available studies of this research question. Additionally, this analysis followed LRA users and non-users from a common index date, using a time-updating treatment indicator to correctly assign time-at-risk to the exposed and unexposed groups. This approach has been less commonly used in the literature and allows for the mitigation of potential immortal time bias. Finally, it is one of the few that provide evidence on the association of lipid-regulating agent use and vascular and other dementias, and used negative and positive controls to assess the potential for residual confounding.

>>>>

In this chapter, I present an analysis of a large population-based electronic health record dataset to investigate the relationship between lipid-regulating agent (LRA) use and dementia outcomes. The analysis aims to address two important limitations of the current evidence base as identified by the systematic review presented in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading).

@@

This second presentation appears to be common in the existing literature on the relationship of statins and dementia, as several of the studies included in the systematic review performed as part of this thesis were identified as being at risk of immortal time bias following formal risk of bias assessment using the ROBINS-I tool (see Section \@ref(risk-of-bias-res)).

@@

#### Impact of dementia code lists {#comparing-codelists}

@@

This section will expand on a potential explanation for the observed results detailed above. However, as the comparison of evidence across different sources is the aim of the triangulation exercise presented in later chapters, a detailed comparison with other published literature will not be provided here, except where needed to illustrate a methodological point. For a detailed comparison of the result presented above with the existing evidence base identified by the systematic review, see Chapter \@ref(tri-heading).

@@

Given the absence of vascular dementia in the published literature, as highlighted in the previous chapter, the unexpected increase in vascular dementia risk with statin use is particularly interesting. It is possible that this finding reflects publication bias if previous research encountered similar methodological issues to this analysis, and these results did not make it into the evidence base as unexpected or assumedly incorrect results are less likely to be submitted or accepted for publication.

@@

The primary strength of this analysis is the relative size of the CPRD and length of follow-up. Having reviewed the existing literature, as identified by the systematic review element of this thesis, this analysis of `r comma(attrition[14,1])` participants is one of the largest available studies of this research question. Additionally, this analysis followed LRA users and non-users from a common index date, using a time-updating treatment indicator to correctly assign time-at-risk to the exposed and unexposed groups. This approach has been less commonly used in the literature and allows for the mitigation of potential immortal time bias. Finally, it is one of the few that provide evidence on the association of lipid-regulating agent use and vascular and other dementias, and used negative and positive controls to assess the potential for residual confounding.

### 2021-12-07 06-Chp-IPD-Analysis.Rmd
Fix broken cross-links

Previous work has suggested a difference in the effect of lipids on dementia risk based on participant age and sex,[@ancelin2013; @mielke2010] though the systematic review presented in Chapter \@ref(sys-rev-heading) did not identify strong evidence for an effect of sex or age at lipid measurement on the relationship between lipids and dementia. However, the number of included studies in the lipid fraction meta-analyses was small, due to both the relatively small number of studies reporting on the lipids exposure and the poor reporting of summary statistics of participant characteristics in those that did. Additionally, best practice guidance recommends against basing the decision to perform an IPD meta-analysis on the between-study heterogeneity in a summary data meta-analysis, as similar distributions of participant covariates across studies may mask a true effect.[@riley2020]

@@

The first approach focused on previously analysed observational prospective cohort studies examining the effect of blood lipid levels on dementia outcomes, as identified by the systematic review presented in Chapter \@ref(sys-rev-heading)  The data sources used in each of these cohort analyses were screened against the criteria listed in the previous section, and eligible cohorts were approached for data access. In the first instance, the first/corresponding author of the publication was emailed (see Appendix \@ref(ipd-email-collab) for a copy of the text and documentation sent to potential collaborators) in Autumn 2020. If this approach did not elicit a response within two months, the last author was contacted, on the basis that the first/corresponding author may have been a more junior member of the research group who had moved to a different institution.

@@

In this analysis, a two-stage IPD analysis was planned. Under a two-stage approach, estimates for the effect of each lipid fraction on incident dementia were first calculated for each data source. More specifically, a logistic regression model adjusted for relevant covariates as detailed in the above section was used to quantify the effect of a 1 _mmol/L_ increase in each lipid fraction on each dementia outcome. Results were expressed as odds ratios (OR). Examination of the data available via the DPUK indicated that the common time-to-event approach used in studies of dementia outcomes would be precluded by the absence of detailed time-to-event data. An overall effect estimate was then produced by combining data-source-specific estimates in a random-effects meta-analysis (see Section \@ref(hold) in Chapter \@ref(hold) for a broader discussion of meta-analysis methods).

@@

Of note, the Whitehall II cohort was analysed in one of the included studies identified by the systematic review presented in Chapter \@ref(sys-rev-heading),[@tynkkynen2018] meaning that a comparison between the published result and the analysis reported here was was possible.

@@

Across the three cohorts, `r n_total_ipd` participants were included in this analysis (Whitehall II = 8208, EPIC = 1115, CaPS = 2512). All cohorts contained data on the four exposure variables of interest (or sufficient data from which to calculate them) and on all-cause dementia outcomes. The only other dementia outcome examined across cohorts was vascular dementia, which was reported in the CaPS and Whitehall II studies. The definitions of dementia outcomes used across cohorts can be seen in Appendix \@ref(appendix-ipd-dementia-def). Cumulatively, there were `r n_ipd_dementia` cases of all-cause dementia, with `r n_ipd_vasdem` further classified as vascular dementia. Summary statistics for each cohort are provided in Table \@ref(tab:cohortSummary-table).

@@

A detailed comparison of the findings presented above with the existing evidence base identified by the systematic review (Chapter \@ref(sys-rev-heading)), is presented as part of the triangulation exercise in Chapter \@ref(tri-heading). As such, this discussion will not provide a detailed comparison of the results of this analysis with other published literature, except to compare between this analysis and previously published results using the same data source.

@@

Finally, this analysis provides new evidence on a previously unexplored outcome (vascular dementia), adding to the extremely small evidence base for this outcome identified by the systematic review presented in Chapters \@ref(sys-rev-method-heading) & \@ref(sys-rev-results-heading). This is relevant even in the previously analysed WhiteHall II cohort, as the previous analysis by Tynkkynen _et al._ did not report on this outcome.

>>>>

Previous work has suggested a difference in the effect of lipids on dementia risk based on participant age and sex,[@ancelin2013; @mielke2010] though the systematic review presented in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading) did not identify strong evidence for an effect of sex or age at lipid measurement on the relationship between lipids and dementia. However, the number of included studies in the lipid fraction meta-analyses was small, due to both the relatively small number of studies reporting on the lipids exposure and the poor reporting of summary statistics of participant characteristics in those that did. Additionally, best practice guidance recommends against basing the decision to perform an IPD meta-analysis on the between-study heterogeneity in a summary data meta-analysis, as similar distributions of participant covariates across studies may mask a true effect.[@riley2020]

@@

The first approach focused on previously analysed observational prospective cohort studies examining the effect of blood lipid levels on dementia outcomes, as identified by the systematic review presented in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading). The data sources used in each of these cohort analyses were screened against the criteria listed in the previous section, and eligible cohorts were approached for data access. In the first instance, the first/corresponding author of the publication was emailed (see Appendix \@ref(ipd-email-collab) for a copy of the text and documentation sent to potential collaborators) in Autumn 2020. If this approach did not elicit a response within two months, the last author was contacted, on the basis that the first/corresponding author may have been a more junior member of the research group who had moved to a different institution.

@@

In this analysis, a two-stage IPD analysis was planned. Under a two-stage approach, estimates for the effect of each lipid fraction on incident dementia were first calculated for each data source. More specifically, a logistic regression model adjusted for relevant covariates as detailed in the above section was used to quantify the effect of a 1 _mmol/L_ increase in each lipid fraction on each dementia outcome. Results were expressed as odds ratios (OR). Examination of the data available via the DPUK indicated that the common time-to-event approach used in studies of dementia outcomes would be precluded by the absence of detailed time-to-event data. An overall effect estimate was then produced by combining data-source-specific estimates in a random-effects meta-analysis (see Section \@ref(meta-analysis-methods) for a broader discussion of meta-analysis methods).

@@

Of note, the Whitehall II cohort was analysed in one of the included studies identified by the systematic review presented in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading),[@tynkkynen2018] meaning that a comparison between the published result and the analysis reported here was was possible.

@@

Across the three cohorts, `r n_total_ipd` participants were included in this analysis (Whitehall II = 8208, EPIC = 1115, CaPS = 2512). All cohorts contained data on the four exposure variables of interest (or sufficient data from which to calculate them) and on all-cause dementia outcomes. The only other dementia outcome examined across cohorts was vascular dementia, which was reported in the CaPS and Whitehall II studies. The definitions of dementia outcomes used across cohorts can be seen in Appendix \@ref(appendix-ipd-dementia-def). Cumulatively, there were `r n_ipd_dementia` cases of all-cause dementia, with `r n_ipd_vasdem` further classified as vascular dementia. Summary statistics for each cohort are provided in Table \@ref(tab:covariateSummary-table).

@@

A detailed comparison of the findings presented above with the existing evidence base identified by the systematic review (Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading)), is presented as part of the triangulation exercise in Chapter \@ref(tri-heading). As such, this discussion will not provide a detailed comparison of the results of this analysis with other published literature, except to compare between this analysis and previously published results using the same data source.

@@

Finally, this analysis provides new evidence on a previously unexplored outcome (vascular dementia), adding to the extremely small evidence base for this outcome identified by the systematic review presented in Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading). This is relevant even in the previously analysed WhiteHall II cohort, as the previous analysis by Tynkkynen _et al._ did not report on this outcome.

### 2021-12-07 07-Chp-Triangulation.Rmd
Fix broken cross-links

Analysis of lipid levels across both the systematic review (Chapter \@ref(sy-rev-results-heading)) and the IPD analysis (Chapter \@ref(ipd-analysis-heading)) both found extremely weak evidence for an association of any lipid fraction with all-cause dementia (TC `r obsLipids$Dementia_TC$estimate`; LDL-c `r obsLipids$Dementia_LDL$estimate`; HDL-c `r obsLipids$Dementia_HDL$estimate`; triglycerides `r obsLipids$Dementia_TG$estimate`). Similarly, meta-analysis of studies identified by the review which examined a binary hypercholesterolemia provided weak evidence for an association of this risk factor with Alzheimer's disease (`r obsHyperchol$Dementia$estimate`). Finally, a meta-analysis of RCTs of statin use (`r rct_statin_acd`) and of MR analyses of genetic instruments that mimic statin use (`r mrStatin$Dementia$estimate`) both found weak evidence of an association, providing indirect evidence on the effect of lipid lowering on this outcome.

@@

To aid with the quantitative triangulation exercise, I created a new method to visualise the extent, direction and type of bias in each result, enabling detailed comparison across different studies contributing to a synthesis. Similar to the standard paired forest plots introduced in Chapter \@ref(sys-rev-), these "bias direction" plots show the level (or extent) of bias across the domains for each result reported using coloured blocks. However, in contrast to the previous forest plots, symbols are used to illustrate the predicted absolute direction of bias in that domain. Distinct symbol pairs are used to denote additive and proportional biases. Illustrations of these are provided in the Section \@ref(quant-tri-results).

@@

From the data sources identified in Section \@ref(tri-data-sources), I identified 18 results relevant to my first causal question of interest. Most (n=17) were identified via the systematic review,`r ldl_ad_citations` while Chapter \@ref(cprd-heading) provided an additional estimate from the CPRD analysis.  The extent, type and direction of predicted bias for each result can be seen in the bias-direction plot presented in Figure \@ref(fig:ldlAdBiasDirection), stratified by study design.

@@

From the data sources identified in Section \@ref(tri-data-sources), I identified 7 results relevant to my causal question of interest: 4 were identified via the systematic review,`r tg_vad_citations` while Chapter \@ref(cprd-heading) provided two additional estimates from unanalysed datasets (CaPS and Whitehall II), and Chapter \@ref(ipd-heading) provides an estimate for the effect of fibrates, a treatment for hypertriglyceridaemia. The extent, type and direction of predicted bias for each result can be seen in the bias-direction plot presented in Figure \@ref(fig:tgVadBiasDirection), stratified by study design.

>>>>

Analysis of lipid levels across both the systematic review (Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading)) and the IPD analysis (Chapter \@ref(ipd-heading)) both found extremely weak evidence for an association of any lipid fraction with all-cause dementia (TC `r obsLipids$Dementia_TC$estimate`; LDL-c `r obsLipids$Dementia_LDL$estimate`; HDL-c `r obsLipids$Dementia_HDL$estimate`; triglycerides `r obsLipids$Dementia_TG$estimate`). Similarly, meta-analysis of studies identified by the review which examined a binary hypercholesterolemia provided weak evidence for an association of this risk factor with Alzheimer's disease (`r obsHyperchol$Dementia$estimate`). Finally, a meta-analysis of RCTs of statin use (`r rct_statin_acd`) and of MR analyses of genetic instruments that mimic statin use (`r mrStatin$Dementia$estimate`) both found weak evidence of an association, providing indirect evidence on the effect of lipid lowering on this outcome.

@@

To aid with the quantitative triangulation exercise, I created a new method to visualise the extent, direction and type of bias in each result, enabling detailed comparison across different studies contributing to a synthesis. Similar to the standard paired forest plots introduced in Chapter \@ref(sys-rev-methods-heading), these "bias direction" plots show the level (or extent) of bias across the domains for each result reported using coloured blocks. However, in contrast to the previous forest plots, symbols are used to illustrate the predicted absolute direction of bias in that domain. Distinct symbol pairs are used to denote additive and proportional biases. Illustrations of these are provided in the Section \@ref(quant-tri-results).

@@

From the data sources identified in Section \@ref(tri-data-sources), I identified 18 results relevant to my first causal question of interest. Most (n=17) were identified via the systematic review,`r ldl_ad_citations` while Chapter \@ref(cprd-analysis-heading) provided an additional estimate from the CPRD analysis.  The extent, type and direction of predicted bias for each result can be seen in the bias-direction plot presented in Figure \@ref(fig:ldlAdBiasDirection), stratified by study design.

@@

From the data sources identified in Section \@ref(tri-data-sources), I identified 7 results relevant to the causal effect of triglycerides at midlife on vascular dementia. Four were identified via the systematic review,`r tg_vad_citations` while Chapter \@ref(cprd-analysis-heading) provided two additional estimates from unanalysed datasets (CaPS and Whitehall II), and Chapter \@ref(ipd-heading) provides an estimate for the effect of fibrates, a treatment for hypertriglyceridaemia. The extent, type and direction of predicted bias for each result can be seen in the bias-direction plot presented in Figure \@ref(fig:tgVadBiasDirection), stratified by study design.

### 2021-12-07 08-Discussion.Rmd
Fix broken cross-links

In __Chapter \@ref(cprd-heading)__, I described an analysis of the association of lipid-regulating agent use and dementia outcomes in the CPRD, a large electronic health record database. I make use of control outcomes to illustrate insufficiently controlled confounding.

@@

Secondly, each analysis presented here makes use of secondary data sources, be it published literature, electronic health records, or existing cohorts. Secondary data can limit the type of analyses performed (for example, in the IPD analysis presented in Chapter \@ref(ipd-analysis-heading), absence of time-to-event data prevented the use of hazard ratios to quantify dementia risk) and the validity of the data if collected for purposes other than research (for example, the accuracy of dementia diagnoses in electronic health records is known to be variable).[@wilkinson2018]

Thirdly, missing evidence was a common limitation across this thesis. In Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading), several studies stated that they did not report the results of an analysis because the results were not significant. This was particularly common among non-randomised studies of exposure examining blood lipids directly. Similarly, an absence of evidence on vascular dementia, potentially due to a publication bias mechanism, limited the analysis of this outcome. Missing evidence was also as an issue in the IPD meta-analysis reported in Chapter \@ref(hold), in the form of a poor response to data access requests, while the absence of empirically-based distributions of bias/indirectness for use with the triangulation framework limited the credibility of the results produced.

@@

The `medrxivr` tool enables programmatic searching of health-related preprints, and future work could incorporate the tool, along with software to search other literature sources, into an automated searching pipeline as part of a living systematic review.[@elliott2014] In addition, the ready access to preprint data afforded by the tool will enable future meta-epidemiological studies to examine factors which may influence eventual publication (e.g. significant result, geographical location/gender/career stage of first author, or some other factor). Finally, the tool will allow future research to assess the impact of peer review and editorial guidelines on manuscripts through comparison of the same paper at two stages (preprinted versus published). This approach was used to explore the impact of editorial polices on open data sharing, as discussed in Chapter \@ref(sys-rev-tools), but could be applied to other aspects of the publication process.

>>>>

In __Chapter \@ref(cprd-analysis-heading)__, I described an analysis of the association of lipid-regulating agent use and dementia outcomes in the CPRD, a large electronic health record database. I make use of control outcomes to illustrate insufficiently controlled confounding.

@@

Secondly, each analysis presented here makes use of secondary data sources, be it published literature, electronic health records, or existing cohorts. Secondary data can limit the type of analyses performed (for example, in the IPD analysis presented in Chapter \@ref(ipd-heading), absence of time-to-event data prevented the use of hazard ratios to quantify dementia risk) and the validity of the data if collected for purposes other than research (for example, the accuracy of dementia diagnoses in electronic health records is known to be variable).[@wilkinson2018]

Thirdly, missing evidence was a common limitation across this thesis. In Chapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading), several studies stated that they did not report the results of an analysis because the results were not significant. This was particularly common among non-randomised studies of exposure examining blood lipids directly. Similarly, an absence of evidence on vascular dementia, potentially due to a publication bias mechanism, limited the analysis of this outcome. Missing evidence was also as an issue in the IPD meta-analysis reported in Chapter \@ref(ipd-heading), in the form of a poor response to data access requests, while the absence of empirically-based distributions of bias/indirectness for use with the triangulation framework limited the credibility of the results produced.

@@

The `medrxivr` tool enables programmatic searching of health-related preprints, and future work could incorporate the tool, along with software to search other literature sources, into an automated searching pipeline as part of a living systematic review.[@elliott2014] In addition, the ready access to preprint data afforded by the tool will enable future meta-epidemiological studies to examine factors which may influence eventual publication (e.g. significant result, geographical location/gender/career stage of first author, or some other factor). Finally, the tool will allow future research to assess the impact of peer review and editorial guidelines on manuscripts through comparison of the same paper at two stages (preprinted versus published). This approach was used to explore the impact of editorial polices on open data sharing, as discussed in Chapter \@ref(sys-rev-tools-heading), but could be applied to other aspects of the publication process.
